Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-12 preliminaries.tex
Handle more action items

Let $\mathcal{B}$ be the Hilbert space corresponding to one qubit. Let $H:\mathcal{B}^{\otimes n}\rightarrow\mathcal{B}^{\otimes n}$ be Hermitian matrices. We use $H\geq0$ to denote $H$ being positive semidefinite. Let $\lambda(H)$ be the smallest eigenvalue of $H$. The ground states of $H$ are the eigenvectors corresponding to $\lambda(H)$. For matrix $H$ and subspace $S$, let $H\big|_S=\Pi_S H \Pi_S$, where $\Pi_S$ is the projector onto the subspace $S$. For a $T$-qubit Hilbert space, let the state $\ket{\widehat{t}}=\ket{1}^{\otimes t}\otimes \ket{0}^{{\otimes (T-t)}}$.
For $n$-qubit states $\rho_1, \rho_2\in\cB^{\otimes n}$,
we write $F(\rho_1, \rho_2)=\left(\tr\sqrt{\sqrt{\rho_1}\rho_2\sqrt{\rho_1}}\right)^2$ for the fidelity between $\rho_1$ and $\rho_2$.
We write $\norm{\rho_1-\rho_2}_{tr}=\frac{1}{2}\norm{\rho_1-\rho_2}_1$ for the trace distance between $\rho_1$ and $\rho_2$.
We have $\norm{\rho_1-\rho_2}_{tr}\leq\sqrt{1-F(\rho_1, \rho_2)}$.
For two densities $f_1$ and $f_2$ over the same finite domain $X$, we write $\norm{f_1-f_2}_{TV}=\frac{1}{2}\sum_{x\in X}\abs{f_1(x)-f_2(x)}$ for their total variation distance.

@@

\begin{definition}
	Let $\Pi=(P, V)$ be a protocol with security parameter $\lambda$ and accuracy parameter $\varepsilon=\frac{1}{\poly(\lambda)}$.
	We say $\Pi$ is a statistically sound protocol for the sampling problem $(D_x)_{x\in\zo^*}$ with completeness error $c$ if for all $x\in\zo^*$:
	\begin{itemize}
		\item At the end of the protocol, $V$ outputs $(d, z)$ where $d\in\set{\Acc, \Rej}$.
		\item Let $(d, z)\leftarrow(P, V)(x, 1^{1/\varepsilon}, 1^\lambda)$, then $d=\Rej$ with probability at most $c(\lambda)$.
		\item For all cheating prover $P^*$, let $(d, z)\leftarrow(P^*, V)(x, 1^{1/\varepsilon}, 1^\lambda)$.
			Define $z_{ideal}$ by
			$$\begin{cases}
				z_{ideal}=\bot & \text{if } d=\Rej\\
				z_{ideal}\leftarrow D_x & \text{if } d=\Acc
			\end{cases}$$
			then $\norm{(d,z)-(d,z_{ideal})}_{TV}\leq\varepsilon$.
	\end{itemize}
\end{definition}

@@

\begin{definition}
	Let $\Pi=(P, V)$ be a protocol with security parameter $\lambda$ and accuracy parameter $\varepsilon=\frac{1}{\poly(\lambda)}$.
	We say $\Pi$ is a computationally sound protocol for the sampling problem $(D_x)_{x\in\zo^*}$ with completeness error $c$ if for all $x\in\zo^*$:
	\begin{itemize}
		\item At the end of the protocol, $V$ outputs $(d, z)$ where $d\in\set{\Acc, \Rej}$.
		\item Let $(d, z)\leftarrow(P, V)(x, 1^{1/\varepsilon}, 1^\lambda)$, then $d=\Rej$ with probability at most $c(\lambda)$.
		\item For all cheating prover $P^*$, let $(d, z)\leftarrow(P^*, V)(x, 1^{1/\varepsilon}, 1^\lambda)$.
			Define $z_{ideal}$ by
			$$\begin{cases}
				z_{ideal}=\bot & \text{if } d=\Rej\\
				z_{ideal}\leftarrow D_x & \text{if } d=\Acc
			\end{cases}$$
			then for all adversary, its probability to successfully distinguish a sample of $(d, z)$ from that of $(d, z_{ideal})$ is less than $\frac{1}{2}+\frac{\varepsilon}{2}$
			\Ethan{Something like this? Or need protocol or specify prior probability is 1/2 each?}
	\end{itemize}
\end{definition}

>>>>

Let $\mathcal{B}$ be the Hilbert space corresponding to one qubit. Let $H:\mathcal{B}^{\otimes n}\rightarrow\mathcal{B}^{\otimes n}$ be Hermitian matrices. We use $H\geq0$ to denote $H$ being positive semidefinite. Let $\lambda(H)$ be the smallest eigenvalue of $H$. The ground states of $H$ are the eigenvectors corresponding to $\lambda(H)$. For matrix $H$ and subspace $S$, let $H\big|_S=\Pi_S H \Pi_S$, where $\Pi_S$ is the projector onto the subspace $S$. For a $T$-qubit Hilbert space, let the state $\ket{\widehat{t}}=\ket{1}^{\otimes t}\otimes \ket{0}^{{\otimes (T-t)}}$.
For $n$-qubit states $\rho_1, \rho_2\in\cB^{\otimes n}$,
we write $F(\rho_1, \rho_2)=\left(\tr\sqrt{\sqrt{\rho_1}\rho_2\sqrt{\rho_1}}\right)^2$ for the fidelity between $\rho_1$ and $\rho_2$.
We write $\norm{\rho_1-\rho_2}_{tr}=\frac{1}{2}\norm{\rho_1-\rho_2}_1$ for the trace distance between $\rho_1$ and $\rho_2$.
We have $\norm{\rho_1-\rho_2}_{tr}\leq\sqrt{1-F(\rho_1, \rho_2)}$.
For two densities $f_1$ and $f_2$ over the same finite domain $X$, we write $\norm{f_1-f_2}_{TV}=\frac{1}{2}\sum_{x\in X}\abs{f_1(x)-f_2(x)}$ for their total variation distance.
We write $x\xleftarrow{\$}X$ when $x$ is sampled uniformly at random from the set $X$.

@@

\begin{definition}
	\label{dfn:stats-secure-proto-sampbqp}
	Let $\Pi=(P, V)$ be a protocol with security parameter $\lambda$ and accuracy parameter $\varepsilon=\frac{1}{\poly(\lambda)}$. \Ethan{Or, where to put this poly dependence? And is this $\exists\poly$ or $\forall\poly$? Both sounds strange.}
	We say $\Pi$ is a statistically sound protocol for the sampling problem $(D_x)_{x\in\zo^*}$ with completeness error $c$ if for all $x\in\zo^*$:
	\begin{itemize}
		\item At the end of the protocol, $V$ outputs $(d, z)$ where $d\in\set{\Acc, \Rej}$.
		\item Let $(d, z)\leftarrow(P, V)(x, 1^{1/\varepsilon}, 1^\lambda)$, then $d=\Rej$ with probability at most $c(\lambda)$.
		\item For all cheating prover $P^*$, let $(d, z)\leftarrow(P^*, V)(x, 1^{1/\varepsilon}, 1^\lambda)$.
			Define $z_{ideal}$ by
			$$\begin{cases}
				z_{ideal}=\bot & \text{if } d=\Rej\\
				z_{ideal}\leftarrow D_x & \text{if } d=\Acc
			\end{cases}$$
			then $\norm{(d,z)-(d,z_{ideal})}_{TV}\leq\varepsilon$.
	\end{itemize}
\end{definition}

@@

\begin{definition}
	Define everything as in \Cref{dfn:stats-secure-proto-sampbqp}.
	We say $\Pi=(P, V)$ is \emph{computationally sound} instead of statistically if instead of 
	$\norm{(d,z)-(d,z_{ideal})}_{TV}\leq\varepsilon$,
	we have that the probability for any adversary to successfully distinguish a sample of $(d, z)$ from that of $(d, z_{ideal})$ is less than $\frac{1}{2}+\frac{\varepsilon}{2}$.
	\Ethan{Something like this? Or need protocol or specify prior probability is 1/2 each?}
\end{definition}

### 2020-04-12 qpip.tex
Handle more action items

	Inputs:
	\begin{itemize}
		\item Classical input $x\in\zo^n$ to the $\SampBQP$ instance
		\item Security parameter $1^\lambda$ where $\lambda\in\bbN$
		\item Accuracy parameter $1^{1/\varepsilon}$ where $\varepsilon=\frac{1}{\poly(\lambda)}$ \Ethan{Just say between 0 and 1; specify poly lambda in thm statement}
	\end{itemize}

	Ingrediants:
	\begin{itemize}
		\item Let $C$ be a quantum circuit consisting of only Hadamard and Toffoli gates, which on input $x$ samples from some $C_x$ such that $\norm{C_x-D_x}\leq\varepsilon$.
		\item Let $T$ be the number of gates in $C$.
		\item Let $C'$ be $C$ padded with $\frac{6T}{\varepsilon}$ identity gates at the end.
		\item Let $H_{C'(x)}$ be the local Hamiltonian instance associated with the computation of $C'(x)$. \Ethan{Cite previous theorem}
		\item Let $\ket{\psi_{C'(x)}}$ be the ground state of $H_{C'(x)}$.
		\item Let $\cVGS$ be the verification algorithm for $H_{C'(x)}$ as defined in \Cref{AlgGroundStateCheck}.
	\end{itemize}

	Protocol:
	\begin{enumerate}
		\item\label{step:qpip1-state-gen} The honest prover prepares $M=505\frac{T^{21}\lambda^2}{\varepsilon^{31}}$ copies of $\ket{\psi_{C'(x)}}$ and sends all of it to the verifier qubit-by-qubit.
		\item\label{step:qpip1-verify} The verifier samples $I\subset[M]$ s.t. $\abs{I}=m$ where $m=\frac{T^{10}\lambda}{\varepsilon^{14}}$, and $k\xleftarrow{\$}[M]\setminus I$. \hannote{$\xleftarrow{\$}$ ?} \Ethan{Sample randomly from. Will add it to preliminary section.}
			For $i$ from $1$ to $M$, it chooses what to do to the $i$-th copy, $\rho_i$, as follows:
		\begin{enumerate}
			\item If $i\in I$, run $y_i\leftarrow\cV_{\GS}(\rho_i)$.
			\item Else if $i=k$, measure the data register \Ethan{Revisit this name} of $\rho_i$ under the standard basis and save the outcome as $z$.
			\item Else, discard $\rho_i$.
		\end{enumerate}
			Let $Y=\sum_{i\in I} y_i$. If $Y>\frac{m}{2}-\kappa m$ where $\kappa=\frac{c\varepsilon^2}{72\left(T+\frac{T}{\varepsilon}\right)^5}$ and $c\in\bbR$ is the constant specified in \Cref{cor:HamCheck}, then the verifier outputs $(\Acc, z)$.
			Else, it outputs $(\Rej, \bot)$.
	\end{enumerate}
\end{protocol}

Note that $\VSamp$ only needs to apply $X$ and $Z$ measurements, and is classical otherwise.
It is simple to check that $\VSamp$ runs in $\poly(n, \lambda, \frac{1}{\epsilon})$ time.
We now show its completeness and soundness.

\begin{thm}
    \label{QPIP1thm}
	$\PiSamp$ has negligible completeness error and is statistically sound.
	\Ethan{Should also say $\lambda$ and $\varepsilon$ are inverse poly related. Or actually, maybe specify that in soundness dfn.}

	\Ethan{Might want to define verifier being efficient (poly-time) there too}
\end{thm}
\begin{prf}
	For completeness, notice $y_i$ are i.i.d. Bernoulli trials with success probability $\frac{1}{2}$.
	So we can apply Chernoff bound (\Cref{thm:Chernoff}) to get
	$$\Prob{Y>\frac{m}{2}-\varepsilon m}\leq\negl(\lambda).$$
	\Ethan{So its rejection chance is...}

	Now we show soundness.
	Suppose $\PSampstar$ is a cheating prover that sends some $\sigma$ to the verifier. \Ethan{on inputs x, eps, lambda...}

@@

	Now we apply de Finetti's theorem (\Cref{deFinetti}) to approximate our measurement result on $\sigma$ with that of a classical mix over $m+1$-fold tensor products of identical copies of some states $\tau_j$.
	That is, there exists some $\rho=\sum_j w_j\tau_j^{\otimes m+1}$
	such that for all quantum-classical channels $\Lambda_i$ acting on registers with size of $\ket{\psi_{C'(x)}}$ \Ethan{Make that size another variable}:
	$$\norm{\Lambda_1\otimes\ldots\otimes\Lambda_{m+1}(\sigma'_{\leq m+1}-\rho)}
	\leq\sqrt{\frac{2m^2\left(n+T+\frac{6T}{\varepsilon}\right)}{M-m}}
	\leq\frac{\varepsilon}{6}$$
	where $\sigma'_{\leq m+1}$ is the first $m+1$ registers of $\sigma'$.
	In our context, for $1\leq i\leq m$, $\Lambda_i$ corresponds to measurements chosen by $\cVGS$.
	$\Lambda_{m+1}$ measures the data register under the standard basis.

>>>>

	Inputs:
	\begin{itemize}
		\item Classical input $x\in\zo^n$ to the $\SampBQP$ instance
		\item Security parameter $1^\lambda$ where $\lambda\in\bbN$
		\item Accuracy parameter $1^{1/\varepsilon}$ where $\varepsilon\in(0, 1)$
	\end{itemize}

	Ingrediants:
	\begin{itemize}
		\item Let $C$ be a quantum circuit consisting of only Hadamard and Toffoli gates, which on input $x$ samples from some $C_x$ such that $\norm{C_x-D_x}\leq\varepsilon$.
		\item Let $T$ be the number of gates in $C$.
		\item Let $C'$ be $C$ padded with $\frac{6T}{\varepsilon}$ identity gates at the end.
		\item Let $H_{C'(x)}$ be the local Hamiltonian instance associated with the computation of $C'(x)$. \Ethan{Cite previous theorem}
		\item Let $\ket{\psi_{C'(x)}}$ be the ground state of $H_{C'(x)}$.
		\item Let $s=n+T+\frac{6T}{\varepsilon}$ be the number of qubits in $\ket{\psi_{C'(x)}}$
		\item Let $\cVGS$ be the verification algorithm for $H_{C'(x)}$ as defined in \Cref{AlgGroundStateCheck}.
	\end{itemize}

	Protocol:
	\begin{enumerate}
		\item\label{step:qpip1-state-gen} The honest prover prepares $M=505\frac{T^{21}\lambda^2}{\varepsilon^{31}}$ copies of $\ket{\psi_{C'(x)}}$ and sends all of it to the verifier qubit-by-qubit.
		\item\label{step:qpip1-verify} The verifier samples $I\subset[M]$ s.t. $\abs{I}=m$ where $m=\frac{T^{10}\lambda}{\varepsilon^{14}}$, and $k\xleftarrow{\$}[M]\setminus I$. \hannote{$\xleftarrow{\$}$ ?} \Ethan{Sample randomly from. Will add it to preliminary section.}
			For $i$ from $1$ to $M$, it chooses what to do to the $i$-th copy, $\rho_i$, as follows:
		\begin{enumerate}
			\item If $i\in I$, run $y_i\leftarrow\cV_{\GS}(\rho_i)$.
			\item Else if $i=k$, measure the data register \Ethan{Revisit this name} of $\rho_i$ under the standard basis and save the outcome as $z$.
			\item Else, discard $\rho_i$.
		\end{enumerate}
			Let $Y=\sum_{i\in I} y_i$. If $Y>\frac{m}{2}-\kappa m$ where $\kappa=\frac{c\varepsilon^2}{72\left(T+\frac{6T}{\varepsilon}\right)^5}$ and $c\in\bbR$ is the constant specified in \Cref{cor:HamCheck}, then the verifier outputs $(\Acc, z)$.
			Else, it outputs $(\Rej, \bot)$.
	\end{enumerate}
\end{protocol}

Note that $\VSamp$ only needs to apply $X$ and $Z$ measurements, and is classical otherwise.
It is simple to check that $\VSamp$ runs in $\poly(n, \lambda, \frac{1}{\epsilon})$ time.
\Ethan{Maybe this is less work than defining a general ``efficient verifier", since $\BQP$ and $\SampBQP$ have different inputs}
We now show its completeness and soundness.

\begin{thm}
    \label{QPIP1thm}
	$\PiSamp$ has negligible completeness error and is statistically sound.
	\Ethan{TODO Fix this. Figure out where to put the inverse poly dependence between lambda and eps.}
\end{thm}
\begin{prf}
	For completeness, notice $y_i$ are i.i.d. Bernoulli trials with success probability $\frac{1}{2}$.
	So we can apply Chernoff bound (\Cref{thm:Chernoff}) to get
	$$\Prob{Y>\frac{m}{2}-\varepsilon m}\leq\negl(\lambda).$$
	So $\VSamp$ has negligible probability to reject.

	Now we show soundness.
	Suppose $\PSampstar$ is a cheating prover that, on inputs $x, 1^{1/\varepsilon}, 1^\lambda$, sends some $\sigma$ to the verifier.

@@

	Now we apply de Finetti's theorem (\Cref{deFinetti}) to approximate our measurement result on $\sigma$ with that of a classical mix over $m+1$-fold tensor products of identical copies of some states $\tau_j$.
	That is, there exists some $\rho=\sum_j w_j\tau_j^{\otimes m+1}$
	such that for all quantum-classical channels $\Lambda_i$ acting on registers of size $s$:
	$$\norm{\Lambda_1\otimes\ldots\otimes\Lambda_{m+1}(\sigma'_{\leq m+1}-\rho)}
	\leq\sqrt{\frac{2m^2\left(n+T+\frac{6T}{\varepsilon}\right)}{M-m}}
	\leq\frac{\varepsilon}{6}$$
	where $\sigma'_{\leq m+1}$ is the first $m+1$ registers of $\sigma'$.
	In our context, for $1\leq i\leq m$, $\Lambda_i$ corresponds to measurements chosen by $\cVGS$.
	$\Lambda_{m+1}$ measures the data register under the standard basis.
